---
title: "Preprocess of drug combination screening data"
author: "Shuyu Zheng"
date: "`r Sys.Date()`"
output: html_notebook
---

```{r}
library(tidyverse)
library(openxlsx)
```

```{r}
output_dir <- "./data/input/drug_combination_raw_experiment_data/processed_response_table/"

if (!dir.exists(output_dir)){
  dir.create(output_dir)
}
```

The drug combination experiments are performed by Jing Jiang. I get the data from her on 2022.11.10. The tables with readouts are saved in the folder "./data/input/drug_combination_raw_experiment_data/2022-11-09/"

This notebook will calculate the dose-response values out from those data and generate a table that is ready for the synergyfinder R package analysis.

There are two designs of the plate layout:

1. Plate 1 to 6: Drug 23, 24, 30 (on dose for each drug) combined with NK cells. NK:target ratio are 0, 1, 2, 4, 6 for plate 9/10 and 0, 2, 4, 8, 12 for other plates.
2. Plate 7, 8: Drug 25, 26 (three dosages for each drug) combined with NK cells. NK:target ratio are 0, 2, 4, 8, 12
3. Plate 11 to 13: Drug 27, 28, 29 (three dosages for each drug) combined with Olaparib. Olaparib dosages are 2uM, 4uM, 7uM and 10uM.


# Match drug ID to drug name

```{r}
drug_name <- data.frame(
  id = c(23, 24, 25, 26, 27, 28, 29, 30),
  drug_name = c(
    "ADU-S100", "STING-IN-2", "Nocodazole", "Clofarabine", "Bortezomib",
    "Carfilzomib", "Pyrrolidinedithiocarbamate ammonium", "H-151"
  )
)
```

```{r}
files <- list.files("./data/input/drug_combination_raw_experiment_data/2022-11-09/")
```

# Combination with NK, plate 1 to 6

```{r}
for (i in 1:6){
  file <- files[startsWith(files, paste0(i, "-"))]
  table <- read.xlsx(
    paste0(
    "./data/input/drug_combination_raw_experiment_data/2022-11-09/", file
    ),
    rows = 15:22, cols = 3:13
  )
  drug1 <- "NK:target"
  drug2_id = strsplit(file, "-")[[1]][5]
  drug2 <- drug_name$drug_name[drug_name$id == drug2_id]

  if (grepl("UWB", file)){
    cell1 <- "UWB1.289"
    cell2 <- "UWB1.289+BRCA1"
    conc2 <- switch(drug2_id,
      "23" = 0.5,
      "24" = 0.25,
      "30" = 0.5
    )
  } else if (grepl("COV", file)){
    cell1 <- "COV362"
    cell2 <- "COV362+BRCA1"
    conc2 <- switch(drug2_id,
      "23" = 1,
      "24" = 0.5,
      "30" = 1
    )
  }

  conc_unit1 <- "ratio"
  conc_unit2 <- "uM"
  
  # The last row is the background signal. We will remove the background signal from whole plates.
  background <- mean(table$`12`)
  table <- table[,1:10] - background
  
  # Negative control
  NC1 <- mean(table[1:3, 5])
  NC2 <- mean(table[1:3, 6])
  
  # Calculate % inhibition
  table[, 1:5] <- 100 * (1 - table[, 1:5]/NC1)
  table[, 6:10] <- 100 * (1 - table[, 6:10]/NC2)
  
  colnames(table) <- c(12, 8, 4, 2, 0, 0, 2, 4, 8, 12)
  df1 <- table[1:3, 1:5] %>% 
    as.data.frame() %>% 
    pivot_longer(everything(), names_to = "conc1", values_to = "response") %>% 
    mutate(
      block_id = 1,
      drug1 = drug1,
      drug2 = drug2,
      cell_line_name = cell1,
      conc2 = 0,
      conc_unit1 = conc_unit1,
      conc_unit2 = conc_unit2
     )
  df2 <- table[4:6, 1:5] %>% 
    as.data.frame() %>% 
    pivot_longer(everything(), names_to = "conc1", values_to = "response") %>% 
    mutate(
      block_id = 1,
      drug1 = drug1,
      drug2 = drug2,
      cell_line_name = cell1,
      conc2 = conc2,
      conc_unit1 = conc_unit1,
      conc_unit2 = conc_unit2
     )
  df3 <- table[1:3, 6:10] %>% 
    as.data.frame() %>% 
    pivot_longer(everything(), names_to = "conc1", values_to = "response") %>% 
    mutate(
      block_id = 2,
      drug1 = drug1,
      drug2 = drug2,
      cell_line_name = cell2,
      conc2 = 0,
      conc_unit1 = conc_unit1,
      conc_unit2 = conc_unit2
     )
  df4 <- table[4:6, 6:10] %>% 
    as.data.frame() %>% 
    pivot_longer(everything(), names_to = "conc1", values_to = "response") %>% 
    mutate(
      block_id = 2,
      drug1 = drug1,
      drug2 = drug2,
      cell_line_name = cell2,
      conc2 = conc2,
      conc_unit1 = conc_unit1,
      conc_unit2 = conc_unit2
     )
  
  df <- Reduce(rbind.data.frame, list(df1, df2, df3, df4))
  df$conc1 <- as.numeric(df$conc1)
  write.csv(df, paste0(output_dir, gsub("xlsx", "csv", file)), row.names = FALSE)
}
```

# Combination with NK, plate 7, 8

```{r}
for (i in 7:8){
  file <- files[startsWith(files, paste0(i, "-"))]
  table <- read.xlsx(
    paste0(
    "./data/input/drug_combination_raw_experiment_data/2022-11-09/", file
    ),
    rows = 15:22, cols = 3:13
  )
  drug1 <- "NK:target"
  drug2_id = strsplit(file, "-")[[1]][3]
  drug2 <- drug_name$drug_name[drug_name$id == drug2_id]
  cell <- "KURAMOCHI"
  conc2 <- switch(drug2_id,
      "25" = c(0, 10, 20, 40),
      "26" = c(0, 1, 3.3, 10),
    )
  conc_unit2 <- switch(drug2_id,
      "25" = "nM",
      "26" = "uM",
    )
  conc_unit1 <- "ratio"
  # The last row is the background signal. We will remove the background signal from whole plates.
  background <- mean(table$`12`)
  table <- table[,1:10] - background
  
  # Negative control
  NC1 <- mean(table[1:3, 5])
  NC2 <- mean(table[1:3, 6])
  
  # Calculate % inhibition
  table[, 1:5] <- 100 * (1 - table[, 1:5]/NC1)
  table[, 6:10] <- 100 * (1 - table[, 6:10]/NC1)
  
  colnames(table) <- c(12, 8, 4, 2, 0, 0, 2, 4, 8, 12)
  df1 <- table[1:3, 1:5] %>% 
    as.data.frame() %>% 
    pivot_longer(everything(), names_to = "conc1", values_to = "response") %>% 
    mutate(
      block_id = 1,
      drug1 = drug1,
      drug2 = drug2,
      cell_line_name = cell,
      conc2 = conc2[1],
      conc_unit1 = conc_unit1,
      conc_unit2 = conc_unit2
     )
  df2 <- table[4:6, 1:5] %>% 
    as.data.frame() %>% 
    pivot_longer(everything(), names_to = "conc1", values_to = "response") %>% 
    mutate(
      block_id = 1,
      drug1 = drug1,
      drug2 = drug2,
      cell_line_name = cell,
      conc2 = conc2[2],
      conc_unit1 = conc_unit1,
      conc_unit2 = conc_unit2
     )
  df3 <- table[1:3, 6:10] %>% 
    as.data.frame() %>% 
    pivot_longer(everything(), names_to = "conc1", values_to = "response") %>% 
    mutate(
      block_id = 1,
      drug1 = drug1,
      drug2 = drug2,
      cell_line_name = cell,
      conc2 = conc2[3],
      conc_unit1 = conc_unit1,
      conc_unit2 = conc_unit2
     )
  df4 <- table[4:6, 6:10] %>% 
    as.data.frame() %>% 
    pivot_longer(everything(), names_to = "conc1", values_to = "response") %>% 
    mutate(
      block_id = 1,
      drug1 = drug1,
      drug2 = drug2,
      cell_line_name = cell,
      conc2 = conc2[4],
      conc_unit1 = conc_unit1,
      conc_unit2 = conc_unit2
     )
  
  df <- Reduce(rbind.data.frame, list(df1, df2, df3, df4))
  df$conc1 <- as.numeric(df$conc1)
  write.csv(df, paste0(output_dir, gsub("xlsx", "csv", file)), row.names = FALSE)
}
```

# Combination with NK, plate 9, 10

```{r}
for (i in 9:10){
  file <- files[startsWith(files, paste0(i, "-"))]
  table <- read.xlsx(
    paste0(
    "./data/input/drug_combination_raw_experiment_data/2022-11-09/", file
    ),
    rows = 15:22, cols = 3:13
  )
  drug1 <- "NK:target"
  drug2_id = strsplit(file, "-")[[1]][5]
  drug2 <- drug_name$drug_name[drug_name$id == drug2_id]

  if (grepl("UWB", file)){
    cell1 <- "UWB1.289"
    cell2 <- "UWB1.289+BRCA1"
    conc2 <- switch(drug2_id,
      "23" = 0.5,
      "24" = 0.25,
      "30" = 0.5
    )
  } else if (grepl("COV", file)){
    cell1 <- "COV362"
    cell2 <- "COV362+BRCA1"
    conc2 <- switch(drug2_id,
      "23" = 1,
      "24" = 0.5,
      "30" = 1
    )
  }

  conc_unit1 <- "ratio"
  conc_unit2 <- "uM"
  
  # The last row is the background signal. We will remove the background signal from whole plates.
  background <- mean(table$`12`)
  table <- table[,1:10] - background
  
  # Negative control
  NC1 <- mean(table[1:3, 5])
  NC2 <- mean(table[1:3, 6])
  
  # Calculate % inhibition
  table[, 1:5] <- 100 * (1 - table[, 1:5]/NC1)
  table[, 6:10] <- 100 * (1 - table[, 6:10]/NC2)
  
  colnames(table) <- c(6, 4, 2, 1, 0, 0, 1, 2, 4, 6)
  df1 <- table[1:3, 1:5] %>% 
    as.data.frame() %>% 
    pivot_longer(everything(), names_to = "conc1", values_to = "response") %>% 
    mutate(
      block_id = 1,
      drug1 = drug1,
      drug2 = drug2,
      cell_line_name = cell1,
      conc2 = 0,
      conc_unit1 = conc_unit1,
      conc_unit2 = conc_unit2
     )
  df2 <- table[4:6, 1:5] %>% 
    as.data.frame() %>% 
    pivot_longer(everything(), names_to = "conc1", values_to = "response") %>% 
    mutate(
      block_id = 1,
      drug1 = drug1,
      drug2 = drug2,
      cell_line_name = cell1,
      conc2 = conc2,
      conc_unit1 = conc_unit1,
      conc_unit2 = conc_unit2
     )
  df3 <- table[1:3, 6:10] %>% 
    as.data.frame() %>% 
    pivot_longer(everything(), names_to = "conc1", values_to = "response") %>% 
    mutate(
      block_id = 2,
      drug1 = drug1,
      drug2 = drug2,
      cell_line_name = cell2,
      conc2 = 0,
      conc_unit1 = conc_unit1,
      conc_unit2 = conc_unit2
     )
  df4 <- table[4:6, 6:10] %>% 
    as.data.frame() %>% 
    pivot_longer(everything(), names_to = "conc1", values_to = "response") %>% 
    mutate(
      block_id = 2,
      drug1 = drug1,
      drug2 = drug2,
      cell_line_name = cell2,
      conc2 = conc2,
      conc_unit1 = conc_unit1,
      conc_unit2 = conc_unit2
     )
  if (i == 10){
    df3 <- filter(df3, conc1 != 6)
    df4 <- filter(df4, conc1 != 6)
  }
  df <- Reduce(rbind.data.frame, list(df1, df2, df3, df4))
  df$conc1 <- as.numeric(df$conc1)
  write.csv(df, paste0(output_dir, gsub("xlsx", "csv", file)), row.names = FALSE)
}
```

# Combination with NK, plate 11 to 13

```{r}
for (i in 11:13){
  file <- files[startsWith(files, paste0(i, "-"))]
  table <- read.xlsx(
    paste0(
    "./data/input/drug_combination_raw_experiment_data/2022-11-09/", file
    ),
    rows = 15:22, cols = 3:13
  )
  drug1 <- "Olaparib"
  drug2_id = strsplit(file, "-")[[1]][3]
  drug2 <- drug_name$drug_name[drug_name$id == drug2_id]
  cell <- "KURAMOCHI"
  conc2 <- switch(drug2_id,
      "27" = c(0, 10, 20, 40),
      "28" = c(0, 2, 4, 8),
      "29" = c(0, 0.5, 1, 2)
    )
  conc_unit2 <- switch(drug2_id,
      "27" = "nM",
      "28" = "nM",
      "29" = "uM"
    )
  conc_unit1 <- "uM"
  # The last row is the background signal. We will remove the background signal from whole plates.
  background <- mean(table$`12`)
  table <- table[,1:10] - background
  
  # Negative control
  NC1 <- mean(table[1:3, 5])
  NC2 <- mean(table[1:3, 6])
  
  # Calculate % inhibition
  table[, 1:5] <- 100 * (1 - table[, 1:5]/NC1)
  table[, 6:10] <- 100 * (1 - table[, 6:10]/NC1)
  
  colnames(table) <- c(10, 7, 4, 2, 0, 0, 2, 4, 7, 10)
  df1 <- table[1:3, 1:5] %>% 
    as.data.frame() %>% 
    pivot_longer(everything(), names_to = "conc1", values_to = "response") %>% 
    mutate(
      block_id = 1,
      drug1 = drug1,
      drug2 = drug2,
      cell_line_name = cell,
      conc2 = conc2[1],
      conc_unit1 = conc_unit1,
      conc_unit2 = conc_unit2
     )
  df2 <- table[4:6, 1:5] %>% 
    as.data.frame() %>% 
    pivot_longer(everything(), names_to = "conc1", values_to = "response") %>% 
    mutate(
      block_id = 1,
      drug1 = drug1,
      drug2 = drug2,
      cell_line_name = cell,
      conc2 = conc2[2],
      conc_unit1 = conc_unit1,
      conc_unit2 = conc_unit2
     )
  df3 <- table[1:3, 6:10] %>% 
    as.data.frame() %>% 
    pivot_longer(everything(), names_to = "conc1", values_to = "response") %>% 
    mutate(
      block_id = 1,
      drug1 = drug1,
      drug2 = drug2,
      cell_line_name = cell,
      conc2 = conc2[3],
      conc_unit1 = conc_unit1,
      conc_unit2 = conc_unit2
     )
  df4 <- table[4:6, 6:10] %>% 
    as.data.frame() %>% 
    pivot_longer(everything(), names_to = "conc1", values_to = "response") %>% 
    mutate(
      block_id = 1,
      drug1 = drug1,
      drug2 = drug2,
      cell_line_name = cell,
      conc2 = conc2[4],
      conc_unit1 = conc_unit1,
      conc_unit2 = conc_unit2
     )
  
  df <- Reduce(rbind.data.frame, list(df1, df2, df3, df4))
  df$conc1 <- as.numeric(df$conc1)
  write.csv(df, paste0(output_dir, gsub("xlsx", "csv", file)), row.names = FALSE)
}
```

# Merge all templates

```{r}
files <- list.files("./data/input/drug_combination_raw_experiment_data/processed_response_table/", full.names = TRUE)

block_id_acc <- 0
df <- NULL
for (f in files){
  print(block_id_acc)
  tmp <- read.csv(f)
  tmp$block_id <- tmp$block_id + block_id_acc
  df <- rbind.data.frame(df, tmp)
  block_id_acc <- max(df$block_id)
}
```

```{r}
write.csv(df, "./data/input/drug_combination_raw_experiment_data/merged_template_2022_11_9.csv", row.names = FALSE)
```

