---
title: "Simulation of ReSisTrace experiment"
output: html_notebook
author: "Shuyu Zheng"
date: "`r Sys.Date()`"
---

```{r}
library(tidyverse)
library(Seurat)
```

# Color code

```{r}
blue = "#3C5488FF"
red =  "#DC0000FF"
green = "#00A087FF"
brown = "#B09C85FF"

blue2 <- "#4DBBD5FF"
red2 <- "#E64B35FF"
green2 <- "#91D1C2FF"
brown2 <- "#7E6148FF"

blue3 <- "#8491B4FF"
red3 <- "#F39B7FFF"
```

# Parameters from KURAMOCHI experiment

## Jun counted during cell culture:

* Total number of cells for experiment at the beginning: 16000
* Cell growth rate after treatments: 2^3 - 2^4
* Cell growth rate without treatment (non-treatment control): 2^6
* Killing rate of treatments: 70%-80%
* Proportion of singletons, twins and quadruplets after doubling:

```{r}
table <- read.csv("./data/output/FigureS1/FigureS1A_cell_division_times.csv")
table
```

```{r}
output_dir <- "./data/output/simulation/"
if (!dir.exists(output_dir)){
  dir.create(output_dir)
}
```

## Parameters from sequencing data:

The number of qualified cells with lineage label from scRNA-seq (drop-out rate):

```{r}
data <- readRDS("./data/input/merged_KURAMOCHI_samples1_scaled_new.RDS")
df <- data@meta.data %>% 
  filter(!is.na(drugSens))
cell_counts <- as.data.frame(table(df$sample, df$time_point))
colnames(cell_counts) <- c("treatment", "time_point", "n_cell")
cell_counts$treatment <- gsub("\\d", "", cell_counts$treatment)
```

The number of cells with lineage label before QC

```{r}
treatments <- c("Carbo", "NK", "Olaparib")
cell_counts <- NULL
for (t in treatments){
  tmp_bt <- nrow(read.csv(paste0("./data/input/preprocess/", ifelse(t == "Carbo", "Carboplatin", t), "_1/bt_lineage_label_table_no_qc.csv")))
  tmp_at <- nrow(read.csv(paste0("./data/input/preprocess/", ifelse(t == "Carbo", "Carboplatin", t), "_1/at_lineage_label_table_no_qc.csv")))
  tmp <- data.frame(
    sample = paste0(t, "_1"),
    treatment = t,
    time_point = "AfterTreatment_no_qc",
    n_cell = tmp_at
  )
  cell_counts <- rbind.data.frame(cell_counts, tmp)
  tmp <- data.frame(
    sample = paste0(t, "_1"),
    treatment = t,
    time_point = "BeforeTreatment_no_qc",
    n_cell = tmp_bt
  )
  cell_counts <- rbind.data.frame(cell_counts, tmp)
  
  tmp_bt <- nrow(read.csv(paste0("./data/input/preprocess/", ifelse(t == "Carbo", "Carboplatin", t), "_2/bt_lineage_label_table_no_qc.csv")))
  tmp_at <- nrow(read.csv(paste0("./data/input/preprocess/", ifelse(t == "Carbo", "Carboplatin", t), "_2/at_lineage_label_table_no_qc.csv")))
  tmp <- data.frame(
    sample = paste0(t, "_2"),
    treatment = t,
    time_point = "AfterTreatment_no_qc",
    n_cell = tmp_at
  )
  cell_counts <- rbind.data.frame(cell_counts, tmp)
  tmp <- data.frame(
    sample = paste0(t, "_2"),
    treatment = t,
    time_point = "BeforeTreatment_no_qc",
    n_cell = tmp_bt
  )
  cell_counts <- rbind.data.frame(cell_counts, tmp)
}

# cell_counts <- cell_counts %>%
#   group_by(treatment, time_point) %>% 
#   summarise(n_cell = mean(n_cell))

write.csv(cell_counts, paste0(output_dir, "cell_counts_for_each_treatment_samples.csv"), row.names = FALSE)
```

```{r}
cell_counts <- read.csv(paste0(output_dir, "cell_counts_for_each_treatment_samples.csv"))
```

# Define the simulation function

```{r}
ReSisTraceSimulation <- function(
    Niter = 100, # number of iterations will be performed
    SampleSize = 16000, # number of cells at the beginning of the experiment
    
    # GrowthRateAT = 2^6, # Cell growth rate after treatments or after cell culture
    LineagePropAfterDoubling = c(42.46, 53.04, 4.5), # Proportion of singletons, twins and quadruplets after doubling:
    NCellBT, # Number of cells in BT sample
    NCellAT, # Number of cells in AT sample
    # NtwinPairs,
    KillRateTotal = 0.8,
    # KillRateR = 1,
    NCellPreSeq = 10000, # Number of cells after treatment and recovering. 
    NCellSeq = 8000, # Number of cells loaed to scRNA-seq machine
    seed = 123
    ){
  PreResistantProp <- 1 - KillRateTotal
  # Containers for results from each iteration
  Nsister.sample <- rep(0, Niter)
  sister.preBT <- rep(0, Niter)
  sister.preBT.false <- rep(0, Niter)
  sister.BT <- rep(0, Niter)
  sister.BT.false <- rep(0, Niter)
  BTAT.labels <- rep(0, Niter)
  unique.label.preAT <- rep(0, Niter)
  unique.label.preBT <- rep(0, Niter)
  label.preBTpreAT <- rep(0, Niter)
  falsepreSLineageFromUnevenSplit <- rep(0, Niter)
  unique.label.AT <- rep(0, Niter)
  unique.label.BT <- rep(0, Niter)
  preR.cell <- rep(0, Niter)
  preS.cell <- rep(0, Niter)
  label.preR <- rep(0, Niter)
  label.preS <- rep(0, Niter)
  trueR.cell <- rep(0, Niter)
  trueS.cell <- rep(0, Niter)
  falseS.cell <- rep(0, Niter)
  falseR.cell <- rep(0, Niter)
  twin.preBT <- rep(0, Niter)
  twin.BT <- rep(0, Niter)
  twin.BT.prop <- rep(0, Niter)
  trueRpreRratio <- rep(0, Niter)
  trueSpreSratio <- rep(0, Niter)
  trueRpreSratio <- rep(0, Niter)
  trueSpreRratio <- rep(0, Niter)
  trueRpreRCell <- rep(0, Niter)
  trueSpreSCell <- rep(0, Niter)
  trueRpreSCell <- rep(0, Niter)
  trueSpreRCell <- rep(0, Niter)
  trueRpreR <- rep(0, Niter)
  trueSpreS <- rep(0, Niter)
  trueRpreS <- rep(0, Niter)
  trueSpreR <- rep(0, Niter)
  PPVCell <- rep(0, Niter)
  NPVCell <- rep(0, Niter)
  FORCell <- rep(0, Niter)
  FDRCell <- rep(0, Niter)
  TPRCell <- rep(0, Niter)
  FPRCell <- rep(0, Niter)
  FNRCell <- rep(0, Niter)
  TNRCell <- rep(0, Niter)
  BACell <- rep(0, Niter)
  PPV <- rep(0, Niter)
  NPV <- rep(0, Niter)
  FOR <- rep(0, Niter)
  FDR <- rep(0, Niter)
  TPR <- rep(0, Niter)
  FPR <- rep(0, Niter)
  FNR <- rep(0, Niter)
  TNR <- rep(0, Niter)
  BA <- rep(0, Niter)
  trueRpreRCellSingleton <- rep(0, Niter)
  trueSpreSCellSingleton <- rep(0, Niter)
  trueRpreSCellSingleton <- rep(0, Niter)
  trueSpreRCellSingleton <- rep(0, Niter)
  trueRpreRSingleton <- rep(0, Niter)
  trueSpreSSingleton <- rep(0, Niter)
  trueRpreSSingleton <- rep(0, Niter)
  trueSpreRSingleton <- rep(0, Niter)
  PPVCellSingleton <- rep(0, Niter)
  NPVCellSingleton <- rep(0, Niter)
  FORCellSingleton <- rep(0, Niter)
  FDRCellSingleton <- rep(0, Niter)
  TPRCellSingleton <- rep(0, Niter)
  FPRCellSingleton <- rep(0, Niter)
  FNRCellSingleton <- rep(0, Niter)
  TNRCellSingleton <- rep(0, Niter)
  BACellSingleton <- rep(0, Niter)
  PPVSingleton <- rep(0, Niter)
  NPVSingleton <- rep(0, Niter)
  FORSingleton <- rep(0, Niter)
  FDRSingleton <- rep(0, Niter)
  TPRSingleton <- rep(0, Niter)
  FPRSingleton <- rep(0, Niter)
  FNRSingleton <- rep(0, Niter)
  TNRSingleton <- rep(0, Niter)
  BASingleton <- rep(0, Niter)
  for(i in 1:Niter){
    cat(i,"\r")
    set.seed(seed + i)
    # Set the initial labeled cell sample (singleton and twins)
    # Nsingleton <- (SampleSize - 2 * NtwinPairs)
    # Label for singletons
    # sample.labels <- c(seq(1, Nsingleton))
    # Append labels for twins
    # if (NtwinPairs > 0){
    #   sample.labels <- c(sample.labels, 
    #                      rep(seq((Nsingleton + 1), (Nsingleton + NtwinPairs)),
    #                          times = 2))
    # }
    sample.labels <- paste0("L", "_", seq(1, SampleSize))
    
    # Label the lineages with Sensitive or Resistant
    Nresistant <- SampleSize * PreResistantProp
    resistant.labels <- sample(unique(sample.labels), Nresistant)
    sensitive.labels <- setdiff(sample.labels, resistant.labels)
    sample.labels <- c(
      paste0(resistant.labels, "_R"),
      paste0(sensitive.labels, "_S")
    )
    
    # # Record the number of sister cells in sample
    # sister.sample <- table(sample.labels)
    # Nsister.sample[i]<- length(sample.labels) - sum(sister.sample == 1)
    # # Label IDs of the sisters in sample (Identifier for false sisters)
    # sister.sample <- names(sister.sample)[sister.sample >= 2]
    # 
    
    # Culture cells for 1x doubling time
    NSingleton <- round(LineagePropAfterDoubling[1] / 100 * SampleSize , 0)
    singleton.lable <- sample(sample.labels, NSingleton)
    NTwin <- round(LineagePropAfterDoubling[2] /100 * SampleSize, 0)
    twin.label <- sample(setdiff(sample.labels, singleton.lable), NTwin)
    NQuadruplet <- SampleSize - NSingleton - NTwin
    quadruplet.label <- setdiff(sample.labels, c(singleton.lable, twin.label))
    
    cellDouble <- seq(1:(NSingleton + 2 * NTwin + 4 * NQuadruplet)) # Unique IDs for each cell after doubling
    cellDouble.label <- c(singleton.lable, rep(twin.label, 2), rep(quadruplet.label, 4)) # Name elements in cells with lineage label ID
    
    # split labeled cells into preBT and preAT
    preBT <- sample(cellDouble, length(cellDouble) / 2, replace = FALSE)
    preAT <- setdiff(cellDouble, preBT)
    preBT.label <- cellDouble.label[preBT]
    preAT.label <- cellDouble.label[preAT]
    unique.label.preBT[i] <- length(unique(preBT.label))
    unique.label.preAT[i] <- length(unique(preAT.label))
    labels.preBTpreAT <- intersect(preBT.label, preAT.label)
    label.preBTpreAT[i] <- length(labels.preBTpreAT)
    label.preATNotPreBT <- setdiff(preAT.label, preBT.label)
    # Number of sister cells in sample(pre-BT)
    sister.preBT[i] <- length(preBT.label) - sum(table(preBT.label) == 1)
    twin.preBT[i] <- sum(table(preBT.label) == 2)
    # # Number of false sisters (1. Label ID are in sister.sample; 2. occurred
    # # more than once)
    # false.sister <- preBT.label[preBT.label %in% sister.sample]
    # sister.preBT.false[i] <- length(false.sister) - sum(table(false.sister) == 1)
    # 
    # Cells passed the scRNA-seq and quality control with lineage label
    if (NCellBT > length(preBT)) {
      warning("NCellBT = ", NCellBT, ". It is larger than the cell number after recovering (", length(preBT), "). The NCellBT is automatically set as ", length(preBT), ".\n")
      NCellBT <- length(preBT)
    }
    BT <- sample(preBT, NCellBT, replace = FALSE)
    BT.label <- cellDouble.label[BT]
    sister.BT[i] <- length(BT.label) - sum(table(BT.label) == 1)
    twin.BT[i] <- sum(table(BT.label) == 2)
    twin.BT.prop[i] <- sum(table(BT.label) == 2)/length(unique(BT.label))
    # # False sisters in BT sample
    # BT.false.sister <- BT.label[BT.label %in% sister.sample]
    # sister.BT.false[i] <- length(BT.false.sister) - sum(table(BT.false.sister) == 1)
    
    # AT treated by drugs and KillRateTotal of cells are died. Resistant cells dead following the KillRateR
    preATR.label <- preAT.label[endsWith(preAT.label, "_R")]
    #preATS.label <-  preAT.label[endsWith(preAT.label, "_S")]
    
    ## Add treatment
    # ATR.label <- sample(preATR.label, length(preATR.label) * (1 - KillRateR), replace = FALSE)
    # NATS <- round(length(preAT) * (1 - KillRateTotal) - length(ATR.label))
    # if (length(preATS.label) <= NATS){
    #   ATS.label <- preATS.label
    # } else {
    #   ATS.label <- sample(preATS.label, NATS)
    # }
    # AT.label.treat <- c(ATR.label, ATS.label)
    
    # After treatment all resistant cells survived and all sensitive
    AT.label.treat <- preATR.label
    
    ## Recover the cells after treatment
    GrowthRateAT <- round(NCellPreSeq / length(AT.label.treat))
    AT.label.recover <- rep(AT.label.treat, GrowthRateAT)
    AT.cell.recover <- seq(1, length(AT.label.recover))
    
    ## Sampling cells for scRNA-seq
    if (NCellSeq > length(AT.label.recover)) {
      warning("NCellSeq = ", NCellSeq, ". It is larger than the cell number after recovering (", length(AT.label.recover), "). The NCellSeq is automatically set as ", length(AT.label.recover), ".\n")
      NCellSeq <- length(AT.label.recover)
    }
    
    AT.cell.seq <- sample(AT.cell.recover, NCellSeq, replace = FALSE)
    
    ## Cells passed the scRNA-seq and quality control with lineage label
    if (NCellAT > NCellSeq) {
      warning("NCellAT = ", NCellAT, ". It is larger than the cell number after recovering (", NCellSeq, "). The NCellAT is automatically set as ", NCellSeq, ".\n")
      NCellAT <- NCellSeq
    }
    AT <- sample(AT.cell.seq, NCellAT, replace = FALSE)
    AT.label <- AT.label.recover[AT]
    
    BTAT.label <- intersect(BT.label, AT.label)
    BTAT.labels[i] <- length(BTAT.label)
    
    unique.label.AT[i] <- length(unique(AT.label)) - length(BTAT.label)
    unique.label.BT[i] <- length(unique(BT.label)) - length(BTAT.label)

    # Number of pre-resistant, pre-sensitive, resistant, new cells
    preR.cell[i] <- sum(BT.label %in% BTAT.label)
    preS.cell[i] <- length(BT.label) - preR.cell[i]
    preR.label <- BT.label[BT.label %in% BTAT.label]
    preS.label <- BT.label[!BT.label %in% BTAT.label]
    label.preR[i] <- length(unique(preR.label))
    label.preS[i] <- length(unique(preS.label))
    
    preR.lineage <- unique(preR.label)
    preS.lineage <- unique(preS.label)
    
    trueRpreR[i] <- sum(endsWith(preR.lineage, "_R")) # TP lineage
    trueSpreS[i] <- sum(endsWith(preS.lineage, "_S")) # TN lineage
    trueRpreS[i] <- sum(endsWith(preS.lineage, "_R")) # FN lineage
    trueSpreR[i] <- sum(endsWith(preR.lineage, "_S")) # FP lineage
    
    trueRpreRCell[i] <- sum(endsWith(preR.label, "_R")) # TP
    trueSpreSCell[i] <- sum(endsWith(preS.label, "_S")) # TN
    trueRpreSCell[i] <- sum(endsWith(preS.label, "_R")) # FN
    trueSpreRCell[i] <- sum(endsWith(preR.label, "_S")) # FP
    falsepreSLineageFromUnevenSplit[i] <- length(intersect(AT.label, label.preATNotPreBT))
    
    # Singletons in BT sample
    singleton.BT.lable <- names(table(BT.label)[table(BT.label) == 1])
    preR.lineage.singleton <- intersect(preR.label, singleton.BT.lable)
    preS.lineage.singleton <- intersect(preS.label, singleton.BT.lable)
    trueRpreRSingleton[i] <- sum(endsWith(preR.lineage.singleton, "_R")) # TP lineage
    trueSpreSSingleton[i] <- sum(endsWith(preS.lineage.singleton, "_S")) # TN lineage
    trueRpreSSingleton[i] <- sum(endsWith(preS.lineage.singleton, "_R")) # FN lineage
    trueSpreRSingleton[i] <- sum(endsWith(preR.lineage.singleton, "_S")) # FP lineage
    
    preR.label.singleton <- preR.label[preR.label %in% singleton.BT.lable]
    preS.label.singleton <- preS.label[preS.label %in% singleton.BT.lable]
    trueRpreRCellSingleton[i] <- sum(endsWith(preR.label.singleton, "_R")) # TP
    trueSpreSCellSingleton[i] <- sum(endsWith(preS.label.singleton, "_S")) # TN
    trueRpreSCellSingleton[i] <- sum(endsWith(preS.label.singleton, "_R")) # FN
    trueSpreRCellSingleton[i] <- sum(endsWith(preR.label.singleton, "_S")) # FP
    
    # Metrices from [contingency table](https://en.wikipedia.org/wiki/Precision_and_recall)
    ## For lineages
    PPV[i] <- trueRpreR[i] / (trueRpreR[i] + trueSpreR[i]) # TP/(TP + FP)
    NPV[i] <- trueSpreS[i] / (trueSpreS[i] + trueRpreS[i]) # TN/(FN + TN)
    FOR[i] <- trueRpreS[i] /(trueSpreS[i] + trueRpreS[i]) # FN/(FN + TN)
    FDR[i] <- trueSpreR[i] / (trueRpreR[i] + trueSpreR[i]) # FP/(TP + FP)
    TPR[i] <- trueRpreR[i] / (trueRpreR[i] + trueRpreS[i]) # TP / (TP + FN)
    FPR[i] <- trueSpreR[i] / (trueSpreR[i] + trueSpreS[i]) # FP / (FP + TN)
    FNR[i] <- trueRpreS[i] / (trueRpreR[i] + trueRpreS[i]) # FN / (TP + FN)
    TNR[i] <- trueSpreS[i] / (trueSpreR[i] + trueSpreS[i]) # TN / (FP + TN)
    BA[i] <-  (TPR[i] + TNR[i]) / 2
    
    ## For cells
    PPVCell[i] <- trueRpreRCell[i] / (trueRpreRCell[i] + trueSpreRCell[i]) # TP/(TP + FP)
    NPVCell[i] <- trueSpreSCell[i] / (trueSpreSCell[i] + trueRpreSCell[i]) # TN/(FN + TN)
    FORCell[i] <- trueRpreSCell[i] /(trueSpreSCell[i] + trueRpreSCell[i]) # FN/(FN + TN)
    FDRCell[i] <- trueSpreRCell[i] / (trueRpreRCell[i] + trueSpreRCell[i]) # FP/(TP + FP)
    TPRCell[i] <- trueRpreRCell[i] / (trueRpreRCell[i] + trueRpreSCell[i]) # TP / (TP + FN)
    FPRCell[i] <- trueSpreRCell[i] / (trueSpreRCell[i] + trueSpreSCell[i]) # FP / (FP + TN)
    FNRCell[i] <- trueRpreSCell[i] / (trueRpreRCell[i] + trueRpreSCell[i]) # FN / (TP + FN)
    TNRCell[i] <- trueSpreSCell[i] / (trueSpreRCell[i] + trueSpreSCell[i]) # TN / (FP + TN)
    BACell[i] <-  (TPRCell[i] + TNRCell[i]) / 2
    
    ## For lineages in singleton
    PPVSingleton[i] <- trueRpreRSingleton[i] / (trueRpreRSingleton[i] + trueSpreRSingleton[i]) # TP/(TP + FP)
    NPVSingleton[i] <- trueSpreSSingleton[i] / (trueSpreSSingleton[i] + trueRpreSSingleton[i]) # TN/(FN + TN)
    FORSingleton[i] <- trueRpreSSingleton[i] /(trueSpreSSingleton[i] + trueRpreSSingleton[i]) # FN/(FN + TN)
    FDRSingleton[i] <- trueSpreRSingleton[i] / (trueRpreRSingleton[i] + trueSpreRSingleton[i]) # FP/(TP + FP)
    TPRSingleton[i] <- trueRpreRSingleton[i] / (trueRpreRSingleton[i] + trueRpreSSingleton[i]) # TP / (TP + FN)
    FPRSingleton[i] <- trueSpreRSingleton[i] / (trueSpreRSingleton[i] + trueSpreSSingleton[i]) # FP / (FP + TN)
    FNRSingleton[i] <- trueRpreSSingleton[i] / (trueRpreRSingleton[i] + trueRpreSSingleton[i]) # FN / (TP + FN)
    TNRSingleton[i] <- trueSpreSSingleton[i] / (trueSpreRSingleton[i] + trueSpreSSingleton[i]) # TN / (FP + TN)
    BASingleton[i] <-  (TPRSingleton[i] + TNRSingleton[i]) / 2
    
    ## For cells
    PPVCellSingleton[i] <- trueRpreRCellSingleton[i] / (trueRpreRCellSingleton[i] + trueSpreRCellSingleton[i]) # TP/(TP + FP)
    NPVCellSingleton[i] <- trueSpreSCellSingleton[i] / (trueSpreSCellSingleton[i] + trueRpreSCellSingleton[i]) # TN/(FN + TN)
    FORCellSingleton[i] <- trueRpreSCellSingleton[i] /(trueSpreSCellSingleton[i] + trueRpreSCellSingleton[i]) # FN/(FN + TN)
    FDRCellSingleton[i] <- trueSpreRCellSingleton[i] / (trueRpreRCellSingleton[i] + trueSpreRCellSingleton[i]) # FP/(TP + FP)
    TPRCellSingleton[i] <- trueRpreRCellSingleton[i] / (trueRpreRCellSingleton[i] + trueRpreSCellSingleton[i]) # TP / (TP + FN)
    FPRCellSingleton[i] <- trueSpreRCellSingleton[i] / (trueSpreRCellSingleton[i] + trueSpreSCellSingleton[i]) # FP / (FP + TN)
    FNRCellSingleton[i] <- trueRpreSCellSingleton[i] / (trueRpreRCellSingleton[i] + trueRpreSCellSingleton[i]) # FN / (TP + FN)
    TNRCellSingleton[i] <- trueSpreSCellSingleton[i] / (trueSpreRCellSingleton[i] + trueSpreSCellSingleton[i]) # TN / (FP + TN)
    BACellSingleton[i] <-  (TPRCellSingleton[i] + TNRCellSingleton[i]) / 2
  }
  tmp <- list(
    # sisterSample = Nsister.sample,
    sisterCellPreBT = sister.preBT,
    twinLineagePreBT = twin.preBT,
    #           falseSisterPreBT = sister.preBT.false,
    sisterCellBT = sister.BT, # falseSisterBT = sister.BT.false,
    twinLineageBT = twin.BT,
    twinLineageBTProp = twin.BT.prop,
    labelPreBTOnly = unique.label.preBT,
    labelPreATOnly = unique.label.preAT,
    labelPreBTPreAT = label.preBTpreAT,
    labelBTOnly = unique.label.BT,
    labelBTAT = BTAT.labels,
    labelATOnly = unique.label.AT,
    falsepreSLineageFromUnevenSplit = falsepreSLineageFromUnevenSplit,
    cellPreR = preR.cell, cellPreS = preS.cell,
    labelPreR = label.preR, labelPreS = label.preS,
    trueRpreR = trueRpreR,
    trueSpreS = trueSpreS,
    trueRpreS = trueRpreS,
    trueSpreR = trueSpreR,
    trueRpreRCell = trueRpreRCell,
    trueSpreSCell = trueSpreSCell,
    trueRpreSCell = trueRpreSCell,
    trueSpreRCell = trueSpreRCell,
    PPV = PPV,
    NPV = NPV,
    FOR = FOR,
    FDR = FDR,
    TPR = TPR,
    FPR = FPR,
    FNR = FNR,
    TNR = TNR,
    BA = BA,
    PPVCell = PPVCell,
    NPVCell = NPVCell,
    FORCell = FORCell,
    FDRCell = FDRCell,
    TPRCell = TPRCell,
    FPRCell = FPRCell,
    FNRCell = FNRCell,
    TNRCell = TNRCell,
    BACell = BACell,
    trueRpreRSingleton = trueRpreRSingleton,
    trueSpreSSingleton = trueSpreSSingleton,
    trueRpreSSingleton = trueRpreSSingleton,
    trueSpreRSingleton = trueSpreRSingleton,
    trueRpreRCellSingleton = trueRpreRCellSingleton,
    trueSpreSCellSingleton = trueSpreSCellSingleton,
    trueRpreSCellSingleton = trueRpreSCellSingleton,
    trueSpreRCellSingleton = trueSpreRCellSingleton,
    PPVSingleton = PPVSingleton,
    NPVSingleton = NPVSingleton,
    FORSingleton = FORSingleton,
    FDRSingleton = FDRSingleton,
    TPRSingleton = TPRSingleton,
    FPRSingleton = FPRSingleton,
    FNRSingleton = FNRSingleton,
    TNRSingleton = TNRSingleton,
    BASingleton = BASingleton,
    PPVCellSingleton = PPVCellSingleton,
    NPVCellSingleton = NPVCellSingleton,
    FORCellSingleton = FORCellSingleton,
    FDRCellSingleton = FDRCellSingleton,
    TPRCellSingleton = TPRCellSingleton,
    FPRCellSingleton = FPRCellSingleton,
    FNRCellSingleton = FNRCellSingleton,
    TNRCellSingleton = TNRCellSingleton,
    BACellSingleton = BACellSingleton)
  return(tmp)
}
```

# Perform the simulation

```{r}
cellSeq_numers <- readxl::read_xlsx("./data/input/simulation/cell_counts_for_each_treatment_samples_AT.xlsx")
colnames(cellSeq_numers)[1] <- "treatment"
cellSeq_numers <- cellSeq_numers %>% 
  group_by(treatment) %>% 
  summarise_all(mean)
cellSeq_numers
```

## Parameter from real data

```{r}
# Fixed parameters
Niter <- 100
SampleSize <- 16000
seed <- 123
LineagePropAfterDoubling <- c(42.46, 53.04, 4.5)

# Parameters changed
KillRateTotal <- seq(0.7, 0.8, by = 0.01)
```

### Simulation

```{r include=FALSE}
treatments <- unique(cell_counts$treatment)
treatments <- treatments[treatments != "ctrl"]
samples <- unique(cell_counts$sample)
res <- NULL
for (i in 1:length(KillRateTotal)) {
  for (s in samples) {
    t <- gsub("\\_\\d", "", s)
    tmp <- ReSisTraceSimulation(
      Niter = Niter,
      SampleSize = SampleSize,
      LineagePropAfterDoubling = LineagePropAfterDoubling,
      NCellBT = round(cell_counts$n_cell[cell_counts$sample == s & cell_counts$time_point == "BeforeTreatment_no_qc"]),
      NCellAT = round(cell_counts$n_cell[cell_counts$sample == s & cell_counts$time_point == "AfterTreatment_no_qc"]),
      KillRateTotal = KillRateTotal[i],
      NCellPreSeq = round(cellSeq_numers$NCellPreSeq[cellSeq_numers$treatment == t]),
      NCellSeq = round(cellSeq_numers$NCellSeq[cellSeq_numers$treatment == t]),
      seed = seed
    )
    ave <- sapply(tmp, mean)
    ave <- as.data.frame(t(ave))
    colnames(ave) <- paste0(colnames(ave), "_mean")
    sd <- sapply(tmp, sd)
    sd <- as.data.frame(t(sd))
    colnames(sd) <- paste0(colnames(sd), "_sd")
    df <- data.frame(
      sample = s, 
      treatment = t,
      KillRateTotal = KillRateTotal[i]
    )
    df <- cbind.data.frame(df, ave, sd)
    res <- rbind.data.frame(res, df)
  }
}
```

```{r}
write.csv(res, paste0(output_dir, "simulation_results.csv"), row.names = FALSE)
```

### Visualization

```{r}
metrics <- c("PPV", "NPV", "FOR", "FDR", "TPR", "FPR", "FNR", "TNR", "BA",
             "PPVCell", "NPVCell", "FORCell", "FDRCell", "TPRCell", "FPRCell", "FNRCell", "TNRCell", "BACell")
res$treatment[res$treatment == "Carbo"] <- "Carboplatin"
res$treatment <- factor(res$treatment, levels = c("Carboplatin", "Olaparib", "NK"))
res$KillRateTotal <- factor(res$KillRateTotal)
for (m in metrics){
  tmp <- res %>% 
    mutate(
      label = paste0(round(res[[paste0(m, "_mean")]] * 100, 2), "+-", round(res[[paste0(m, "_sd")]] * 100, 2))
    )
  p <- ggplot(
    tmp,
    aes_string(
      x = "treatment", y = "KillRateTotal", fill = paste0(m, "_mean"),
      label = "label"
    )
  ) +
    geom_tile() +
    geom_text(size = 3) +
    scale_fill_gradient2(low = "yellow", high = "red") +
    labs(
      title = ifelse(endsWith(m, "Cell"), "Calculated from cell numbers", "Calculated from lineage numbers"),
      x = "",
      y = "Killing rate on pre-resistant cells",
      fill = switch (gsub("Cell", "", m),
        "FOR" = "False Omission Rate\n(FOR)",
        "PPV" = "Positive predictive value\n(PPV)",
        "NPV" = "Negative predictive value\n(NPV)",
        "FDR" = "False discovery rate\n(FNR)",
        "TPR" = "True positive rate\n(TPR)",
        "FPR" = "False positive rate\n(FPR)",
        "FNR" = "False negative rate\n(FNR)",
        "TNR" = "True negative rate\n(TNR)",
        "BA" = "Balanced accuracy\n(BA)"
      )
    ) +
    theme(
      panel.background = element_blank(),
      legend.title = element_text(size = 8),
      plot.title = element_text(hjust = 0.5),
      plot.subtitle = element_text(hjust = 0.5)
      )
  ggsave(paste0(output_dir, m, "_observed.png"), p, height = 3.5, width = 4.5)
}

```

## Perfect doubling

```{r}
# Fixed parameters
Niter <- 100
SampleSize <- 16000
seed <- 123
LineagePropAfterDoubling <- c(0, 100, 0)

# Parameters changed
KillRateTotal <- seq(0.7, 0.8, by = 0.01)
```

### Simulation

```{r include=FALSE}
treatments <- unique(cell_counts$treatment)
treatments <- treatments[treatments != "ctrl"]
res <- NULL
for (i in 1:length(KillRateTotal)) {
  for (t in treatments) {
    tmp <- ReSisTraceSimulation(
      Niter = Niter,
      SampleSize = SampleSize,
      LineagePropAfterDoubling = LineagePropAfterDoubling,
      NCellBT = round(cell_counts$n_cell[cell_counts$treatment == t & cell_counts$time_point == "BeforeTreatment_no_qc"]),
      NCellAT = round(cell_counts$n_cell[cell_counts$treatment == t & cell_counts$time_point == "AfterTreatment_no_qc"]),
      KillRateTotal = KillRateTotal[i],
      NCellPreSeq = round(cellSeq_numers$NCellPreSeq[cellSeq_numers$treatment == t]),
      NCellSeq = round(cellSeq_numers$NCellSeq[cellSeq_numers$treatment == t]),
      seed = seed
    )
    ave <- sapply(tmp, mean)
    ave <- as.data.frame(t(ave))
    colnames(ave) <- paste0(colnames(ave), "_mean")
    sd <- sapply(tmp, sd)
    sd <- as.data.frame(t(sd))
    colnames(sd) <- paste0(colnames(sd), "_sd")
    df <- data.frame(
      treatment = t,
      KillRateTotal = KillRateTotal[i]
    )
    df <- cbind.data.frame(df, ave, sd)
    res <- rbind.data.frame(res, df)
  }
}
```

```{r}
write.csv(res, paste0(output_dir, "simulation_results_perfect_doubling.csv"), row.names = FALSE)
```

### Visualization

```{r}
metrics <- c("PPV", "NPV", "FOR", "FDR", "TPR", "FPR", "FNR", "TNR", "BA",
             "PPVCell", "NPVCell", "FORCell", "FDRCell", "TPRCell", "FPRCell", "FNRCell", "TNRCell", "BACell")
res$treatment[res$treatment == "Carbo"] <- "Carboplatin"
res$treatment <- factor(res$treatment, levels = c("Carboplatin", "Olaparib", "NK"))
res$KillRateTotal <- factor(res$KillRateTotal)
for (m in metrics){
  tmp <- res %>% 
    mutate(
      label = paste0(round(res[[paste0(m, "_mean")]] * 100, 2), "+-", round(res[[paste0(m, "_sd")]] * 100, 2))
    )
  p <- ggplot(
    tmp,
    aes_string(
      x = "treatment", y = "KillRateTotal", fill = paste0(m, "_mean"),
      label = "label"
    )
  ) +
    geom_tile() +
    geom_text(size = 3) +
    scale_fill_gradient2(low = "yellow", high = "red") +
    labs(
      title = ifelse(endsWith(m, "Cell"), "Calculated from cell numbers", "Calculated from lineage numbers"),
      x = "",
      y = "Killing rate on pre-resistant cells",
      fill = switch (gsub("Cell", "", m),
        "FOR" = "False Omission Rate\n(FOR)",
        "PPV" = "Positive predictive value\n(PPV)",
        "NPV" = "Negative predictive value\n(NPV)",
        "FDR" = "False discovery rate\n(FNR)",
        "TPR" = "True positive rate\n(TPR)",
        "FPR" = "False positive rate\n(FPR)",
        "FNR" = "False negative rate\n(FNR)",
        "TNR" = "True negative rate\n(TNR)",
        "BA" = "Balanced accuracy\n(BA)"
      )
    ) +
    theme(
      panel.background = element_blank(),
      legend.title = element_text(size = 8),
      plot.title = element_text(hjust = 0.5),
      plot.subtitle = element_text(hjust = 0.5)
      )
  ggsave(paste0(output_dir, m, "_perfect_doubling.png"), p, height = 3.5, width = 4.5)
}
```

## No drop out perfect doubling

```{r}
# Fixed parameters
Niter <- 100
SampleSize <- 16000
seed <- 123
LineagePropAfterDoubling <- c(0, 100, 0)
NCellAT <- 16000
NCellBT <- 16000
NCellPreSeq <- 16000
NCellSeq <- 16000

# Parameters changed
KillRateTotal <- seq(0.7, 0.8, by = 0.01)
```

### Simulation

```{r include=FALSE}
res <- NULL
for (i in 1:length(KillRateTotal)) {
  tmp <- ReSisTraceSimulation(
    Niter = Niter,
    SampleSize = SampleSize,
    LineagePropAfterDoubling = LineagePropAfterDoubling,
    NCellBT = NCellBT,
    NCellAT = NCellAT,
    KillRateTotal = KillRateTotal[i],
    NCellPreSeq = NCellPreSeq,
    NCellSeq = NCellSeq,
    seed = seed
  )
  ave <- sapply(tmp, mean)
  ave <- as.data.frame(t(ave))
  colnames(ave) <- paste0(colnames(ave), "_mean")
  sd <- sapply(tmp, sd)
  sd <- as.data.frame(t(sd))
  colnames(sd) <- paste0(colnames(sd), "_sd")
  df <- data.frame(
    KillRateTotal = KillRateTotal[i]
  )
  df <- cbind.data.frame(df, ave, sd)
  res <- rbind.data.frame(res, df)
}

```

```{r}
write.csv(res, paste0(output_dir, "simulation_results_perfect_doubling_no_drop_out.csv"), row.names = FALSE)
```

## No drop out

```{r}
# Fixed parameters
Niter <- 100
SampleSize <- 16000
seed <- 123
LineagePropAfterDoubling <- c(42.46, 53.04, 4.5)
NCellAT <- 13323
NCellBT <- 13323
NCellPreSeq <- 13323
NCellSeq <- 13323

# Parameters changed
KillRateTotal <- seq(0.7, 0.8, by = 0.01)
```

### Simulation

```{r include=FALSE}
res <- NULL
for (i in 1:length(KillRateTotal)) {
  tmp <- ReSisTraceSimulation(
    Niter = Niter,
    SampleSize = SampleSize,
    LineagePropAfterDoubling = LineagePropAfterDoubling,
    NCellBT = NCellBT,
    NCellAT = NCellAT,
    KillRateTotal = KillRateTotal[i],
    NCellPreSeq = NCellPreSeq,
    NCellSeq = NCellSeq,
    seed = seed
  )
  ave <- sapply(tmp, mean)
  ave <- as.data.frame(t(ave))
  colnames(ave) <- paste0(colnames(ave), "_mean")
  sd <- sapply(tmp, sd)
  sd <- as.data.frame(t(sd))
  colnames(sd) <- paste0(colnames(sd), "_sd")
  df <- data.frame(
    KillRateTotal = KillRateTotal[i]
  )
  df <- cbind.data.frame(df, ave, sd)
  res <- rbind.data.frame(res, df)
}

```

```{r}
write.csv(res, paste0(output_dir, "simulation_results_no_drop_out.csv"), row.names = FALSE)
```

# Line plot

```{r}
ideal_condition <- read.csv(paste0(output_dir, "simulation_results_perfect_doubling_no_drop_out.csv"))
observed_condition <- read.csv(paste0(output_dir, "simulation_results.csv"))
plot_table <- ideal_condition %>% 
  mutate(treatment = "Base line") %>% 
  rbind.data.frame(observed_condition)
plot_table$treatment[plot_table$treatment == "Carbo"] <- "Carboplatin"
plot_table$treatment <- factor(plot_table$treatment, levels = c("Carboplatin", "Olaparib", "NK", "Base line"))
plot_table$KillRateTotal <- plot_table$KillRateTotal * 100
```

```{r}
metrics <- c("NPV", "FOR")
for (m in metrics) {
  tmp <- plot_table %>% 
    select(KillRateTotal, treatment, mean = paste0(m, "_mean"), sd = paste0(m, "_sd"))
  tmp$mean <- tmp$mean * 100
  tmp$sd <- tmp$sd * 100
  p<- ggplot(
      tmp,
      aes(
        x = KillRateTotal,
        y = mean,
        group = treatment,
        color = treatment
      )
    ) + 
    geom_line() +
    # geom_point() +
    geom_errorbar(
      aes(ymin = mean - sd, ymax = mean + sd), width=.3,
      position=position_dodge(0.05)
    ) +
    theme_classic() + 
    theme(
      text = element_text(size = 21),
      axis.title = element_text(size = 21, color = "black", hjust = 0.5),
      axis.text = element_text(size = 21, color = "black"),
      legend.text = element_text(size = 21, color = "black")
    ) +
    scale_color_manual(values = c(blue, red, brown, "#000000")) +
    ylim(c(0, 100)) +
    labs(
      x = "Killing rate (%)",
      y = ifelse(m == "NPV", "Pre-sensitive predictive rate (%)", m),
      color = ""
    )
  # ggsave(paste0(output_dir, m, "_lineage_count.pdf"), p, width = 6, height = 3)
  ggsave(paste0(output_dir, m, "_lineage_count.pdf"), p, width = 7, height = 4)
  
  tmp <- plot_table %>% 
    select(KillRateTotal, treatment, mean = paste0(m, "Cell_mean"), sd = paste0(m, "Cell_sd"))
  tmp$mean <- tmp$mean * 100
  tmp$sd <- tmp$sd * 100
  p<- ggplot(
      tmp,
      aes(
        x = KillRateTotal,
        y = mean,
        group = treatment,
        color = treatment
      )
    ) + 
    geom_line() +
    # geom_point() +
    geom_errorbar(
      aes(ymin = mean - sd, ymax = mean + sd), width=.3,
      position=position_dodge(0.05)
    ) +
    theme_classic() + 
    theme(
      text = element_text(size = 21),
      axis.title = element_text(size = 21, color = "black", hjust = 0.5),
      axis.text = element_text(size = 21, color = "black"),
      legend.text = element_text(size = 21, color = "black")
    ) +
    scale_color_manual(values = c(blue, red, brown, "#000000")) +
    ylim(c(0, 100)) +
    labs(
      x = "Killing rate (%)",
      y = ifelse(m == "NPV", "Pre-sensitive predictive rate (%)", m),
      color = ""
    )
  # ggsave(paste0(output_dir, m, "_cell_count.pdf"), p, width = 6, height = 3)
  ggsave(paste0(output_dir, m, "_cell_count.pdf"), p, width = 7, height = 4)
  
  tmp <- plot_table %>% 
    select(KillRateTotal, treatment, mean = paste0(m, "Singleton_mean"), sd = paste0(m, "Singleton_sd"))
  tmp$mean <- tmp$mean * 100
  tmp$sd <- tmp$sd * 100
  p<- ggplot(
      tmp,
      aes(
        x = KillRateTotal,
        y = mean,
        group = treatment,
        color = treatment
      )
    ) + 
    geom_line() +
    # geom_point() +
    geom_errorbar(
      aes(ymin = mean - sd, ymax = mean + sd), width=.3,
      position=position_dodge(0.05)
    ) +
    theme_classic() + 
    theme(
      text = element_text(size = 21),
      axis.title = element_text(size = 21, color = "black", hjust = 0.5),
      axis.text = element_text(size = 21, color = "black"),
      legend.text = element_text(size = 21, color = "black")
    ) +
    scale_color_manual(values = c(blue, red, brown, "#000000")) +
    ylim(c(0, 100)) +
    labs(
      x = "Killing rate (%)",
      y = ifelse(m == "NPV", "Pre-sensitive predictive rate (%)", m),
      color = ""
    )
  # ggsave(paste0(output_dir, m, "_lineage_count_singleton.pdf"), p, width = 6, height = 3)
  ggsave(paste0(output_dir, m, "_lineage_count_singleton.pdf"), p, width = 7, height = 4)
  
  tmp <- plot_table %>% 
    select(KillRateTotal, treatment, mean = paste0(m, "CellSingleton_mean"), sd = paste0(m, "CellSingleton_sd"))
  tmp$mean <- tmp$mean * 100
  tmp$sd <- tmp$sd * 100
  p<- ggplot(
      tmp,
      aes(
        x = KillRateTotal,
        y = mean,
        group = treatment,
        color = treatment
      )
    ) + 
    geom_line() +
    # geom_point() +
    geom_errorbar(
      aes(ymin = mean - sd, ymax = mean + sd), width=.3,
      position=position_dodge(0.05)
    ) +
    theme_classic() + 
    theme(
      text = element_text(size = 21),
      axis.title = element_text(size = 21, color = "black", hjust = 0.5),
      axis.text = element_text(size = 21, color = "black"),
      legend.text = element_text(size = 21, color = "black")
    ) +
    scale_color_manual(values = c(blue, red, brown, "#000000")) +
    ylim(c(0, 100)) +
    labs(
      x = "Killing rate",
      y = ifelse(m == "NPV", "Pre-sensitive predictive rate (%)", m),
      color = ""
    )
  # ggsave(paste0(output_dir, m, "_cell_count_singleton.pdf"), p, width = 6, height = 3)
  ggsave(paste0(output_dir, m, "_cell_count_singleton.pdf"), p, width = 7, height = 4)
}
```
```{r}
p<- ggplot(plot_table, aes(x=KillRateTotal, y=NPVCell_mean * 100, group=treatment, color=treatment)) + 
  geom_line() +
  geom_point() +
  theme_bw() + 
  theme(
    text = element_text(size = 20),
    axis.title = element_text(size = 20, color = "black"),
    axis.text = element_text(size = 20, color = "black"),
    legend.text = element_text(size = 20, color = "black")
  ) +
  scale_color_manual(values = c(blue, red, brown, "#000000")) +
  ylim(c(0, 100)) +
  labs(
    title = "Cell count",
    x = "Killing rate",
    y = "Negative predictive value",
    color = "")
p
```

```{r}
p<- ggplot(plot_table, aes(x=KillRateTotal, y=NPVSingleton_mean * 100, group=treatment, color=treatment)) + 
  geom_line() +
  geom_point() +
  theme_bw() + 
  theme(
    text = element_text(size = 20),
    axis.title = element_text(size = 20, color = "black"),
    axis.text = element_text(size = 20, color = "black"),
    legend.text = element_text(size = 20, color = "black")
  ) +
  scale_color_manual(values = c(blue, red, brown, "#000000")) +
  ylim(c(0, 100)) +
  labs(
    title = "Lineage count (only singletons)",
    x = "Killing rate",
    y = "Negative predictive value",
    color = "")
p
```