---
title: "All before treatment samples"
author: "Shuyu Zheng"
date: "`r Sys.Date()`"
output: 
  html_notebook:
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r include=FALSE}
library(Seurat)
library(ggpubr)
library(tidyverse)
library(wordspace)
options(bitmapType='cairo')
output_dir <- "../data/output/KURAMOCHI_before_treatment/"
if (!dir.exists(output_dir)){
  dir.create(output_dir, recursive = TRUE)
}
```

The input data of this section is the Seurat object for **before treatment sample from all experiments** after quality control and normalization. It is one of the "Preprocess_xx.Rmd" outputs.

# 1. All BT samples

## 1. Load data

```{r eval = FALSE}
merge <- NULL
gene_meta <- NULL
for (i in c("NK1", "NK2", "Carbo1", "Carbo2", "Olaparib1", "Olaparib2", "ctrl1", "ctrl2")){
  tmp <- readRDS(paste0("../data/raw/ReSisTrace_KURAMOCHI_samples/", i, ".RDS"))
  tmp$sample <- rep(i, nrow(tmp@meta.data))
  tmp <- RenameCells(tmp, add.cell.id = i)
  gene_meta <- gene_meta %>%
    rbind.data.frame(tmp@misc$gene_meta) %>%
    unique()
  if (is.null(merge)){
    merge <- tmp
  } else {
    merge <- merge(merge, tmp)
  }
}
merge@misc$gene_meta <- gene_meta
```

## 2. Filter genes lower expressed

Set filtering threshold

```{r}
cutoff_quantile <- 0.1 # 10% quantile for sum expression across all bt samples
exp_sum <- Matrix::rowSums(merge@assays$RNA@data)
exp_mean <- Matrix::rowMeans(merge@assays$RNA@data)
low_exp_genes <- names(exp_sum[exp_sum < quantile(exp_sum, cutoff_quantile)])

cat("With the quantile:", cutoff_quantile * 100, "%.",
    "The cutoff of expression (normalized) is ",
    quantile(exp_sum, cutoff_quantile),
    ". There are ", length(low_exp_genes),
    "genes out of", nrow(merge), "will be reomved.")
```

```{r}
write.csv(
  data.frame(
    gene = low_exp_genes, 
    sum_expression = exp_sum[exp_sum < quantile(exp_sum, cutoff_quantile)],
    stringsAsFactors = FALSE
  ),
  paste0(output_dir, "filtered_low_exp_genes.csv"),
  row.names = FALSE
)

write.csv(
  data.frame(
    gene = names(exp_sum), 
    sum_expression = exp_sum,
    mean_expression = exp_mean,
    stringsAsFactors = FALSE
  ),
  paste0(output_dir, "genes_exp_sum_across_bt.csv"),
  row.names = FALSE
)
```

Filter the low expressed genes

```{r}
merge <- subset(merge, features = rownames(merge)[!rownames(merge) %in% low_exp_genes])
```

## 3. Calculate

```{r}
merge <- FindVariableFeatures(merge)
merge <- ScaleData(merge)
merge <- RunPCA(merge, veverbose = FALSE)
merge <- RunUMAP(merge, dims = 1:20, veverbose = FALSE)
merge <- RunTSNE(merge, dims = 1:20, veverbose = FALSE)
```

```{r}
Idents(merge) <- "sample"
pca <- PCAPlot(merge)
umap <- UMAPPlot(merge)
tsne <- TSNEPlot(merge)
pca
umap
tsne
```
```{r}
ggsave(paste0(output_dir, "/all_bt_sample_filter_gene_pca.pdf"), plot = pca, device = "pdf")
ggsave(paste0(output_dir, "/all_bt_sample_filter_gene_umap.pdf"), plot = umap, device = "pdf")
ggsave(paste0(output_dir, "/all_bt_sample_filter_gene_tsne.pdf"), plot = tsne, device = "pdf")
```

```{r}
saveRDS(merge, file = paste0(output_dir, "/all_bt_sample_merge_filtered_gene.RDS"))
```
```{r eval=FALSE}
merge <- readRDS(paste0(output_dir, "/all_bt_sample_merge_filtered_gene.RDS"))
```
```{r}
merge$treatment <- sub("_.", "", merge$sample)
pca <- PCAPlot(subset(merge, cells = colnames(merge)[!is.na(merge$drugSens)]), split.by = "drugSens", group.by = "treatment")
umap <- UMAPPlot(subset(merge, cells = colnames(merge)[!is.na(merge$drugSens)]), split.by = "drugSens", group.by = "treatment")
tsne <- TSNEPlot(subset(merge, cells = colnames(merge)[!is.na(merge$drugSens)]), split.by = "drugSens", group.by = "treatment")
pca
umap
tsne
```

```{r}
ggsave(paste0(output_dir, "/all_bt_sample_filter_gene_split_by_drugSens_pca.pdf"), plot = pca, device = "pdf")
ggsave(paste0(output_dir, "/all_bt_sample_filter_gene_split_by_drugSens_umap.pdf"), plot = umap, device = "pdf")
ggsave(paste0(output_dir, "/all_bt_sample_filter_gene_split_by_drugSens_tsne.pdf"), plot = tsne, device = "pdf")
```
# 2. Sister Similarity

```{r}
rm(list = setdiff(ls(), c("output_dir", "merge")))
source("./auxiliary_codes.R")
```

## 1. Without filtering sisters
### 1. Plot random sisters

Relabel sisters with the info of sample
```{r}
merge$sisters_sample <- paste(merge$sample, merge$sisters, sep = "_")
```

```{r}
all_sister_pairs <- merge$sisters_sample[!endsWith(merge$sisters_sample, "NA")]
all_sister_pairs <- unique(all_sister_pairs)
cat("There are totally", length(all_sister_pairs),
    "unique sisterhoods across all BT samples")
```

Randomly picking 20 sister groups to plot

```{r}
set.seed(1)
n <- 10
groups <- sample(all_sister_pairs, n)
highlight_group <- vector("list", length(groups))
names(highlight_group) <- groups
for (i in groups){
  highlight_group[[i]] <- colnames(merge)[which(merge$sisters_sample == i)]
}

# myCol <- ggsci::pal_aaas()(n)
myCol <- c('#88CCEE', '#44AA99', '#117733', '#332288', '#DDCC77',
           '#999933', '#CC6677', '#882255', '#AA4499', "#000000")
names(myCol) <- groups
pca <- PCAPlot(merge, cells.highlight = highlight_group,
               cols.highlight = myCol[groups], sizes.highlight = 4) + NoLegend()


umap <- UMAPPlot(merge, cells.highlight = highlight_group,
               cols.highlight = myCol[groups], sizes.highlight = 4) + NoLegend()

tsne <- TSNEPlot(merge, cells.highlight = highlight_group,
               cols.highlight = myCol[groups], sizes.highlight = 4) + NoLegend()

pca
umap
tsne
```

```{r}
  ggsave(paste0(output_dir, "/highlight_sister_group_",paste0(groups, collapse = "."), "_pca.pdf"), plot = pca, device = "pdf")
  ggsave(paste0(output_dir, "/highlight_sister_group_",paste0(groups, collapse = "."), "_umap.pdf"), plot = umap, device = "pdf")
  ggsave(paste0(output_dir, "/highlight_sister_group_",paste0(groups, collapse = "."), "_tsne.pdf"), plot = tsne, device = "pdf")
```

### 2. Similarity analysis

1. Prepare expression matrix

```{r}
expression_mat <- as.matrix(GetAssayData(merge, slot = "data"))
```

2. Prepare sister table (including all sisters)

```{r}
sister_table <- data.frame(cell_barcode = colnames(merge), 
                     sisters = merge$sisters_sample,
                     stringsAsFactors = FALSE) %>% 
  dplyr::filter(sisters %in% all_sister_pairs) # remove singletons

sisters <- NULL
for (i in unique(sister_table$sisters)) {
  tmp_set <- sister_table$cell_barcode[which(sister_table$sisters  == i)]
  tmp_set <- combn(tmp_set, 2)
  tmp_df <- data.frame(group = rep(i, ncol(tmp_set)),
                       sister1 = tmp_set[1, ],
                       sister2 = tmp_set[2, ],
                       stringsAsFactors = FALSE)
  sisters <- rbind.data.frame(sisters, tmp_df)
  tmp_df <- NULL
  tmp_set <- NULL
}
```

```{r}
rm(merge) # remove large seurat object to release memory
rm(pca, tsne, umap)
```

3. Statistic analysis 

We can use function `CalculateSisterSimilarity` (defined in 'script/Auxiliary_codes.R" file) to generate analysis sister similarities by setting the parameters:

* **expression_mat**: the gene expression matrix in log2(normalized_count + 1) scale containing gene names as rows and cell barcodes as columns.
* **n_gene**: number of top variable genes used to calculate the similarity.
* **sister_table**: a data frame containing the group number of sister lineages and corresponding sisters' cell barcodes.

Following is an example using Euclidean distance and top 1000 variable genes.

1. Calculate difference matrix

**Note** This function will take a long time to calculate the difference matrix. Some times it takes hours when there are a lot genes and cells.

```{r}
# time_stamp <- Sys.time()
plot_data <- CalculateSisterSimilarity(
  expression_mat = expression_mat,
  n_gene = 1000,
  sister_table = sisters,
  similarity = "euclidean"
  )
# Sys.time() - time_stamp

write_rds(plot_data, paste0(output_dir, "/sister_distance.RDS"))
```

t-test

```{r}
plot_data <- read_rds(paste0(output_dir, "/sister_distance.RDS"))
```

```{r}
sister_d <- data.frame(
    sister_id = names(plot_data$sister_distance),
    distance = plot_data$sister_distance,
    stringsAsFactors = FALSE
    )
write.csv(
  sister_d,
  paste0(output_dir, "/sister_distance.csv"), row.names = FALSE)
```

```{r}
rm(expression_mat)
```

## 2. Filter sisters far away to each other

```{r}
table(table(sister_d$sister_id))
```

```{r}
cutoff_quantile <- 0.9 # 90% quantile for sister distance across all bt samples
cutoff <- quantile(sister_d$distance, cutoff_quantile)
false_sisters <- unique(sister_d$sister_id[sister_d$distance > cutoff])

sister_d$true_sisters <- !sister_d$sister_id %in% false_sisters
cat("With the quantile:", cutoff_quantile * 100, "%.",
    "The cutoff of distance is",
    quantile(sister_d$distance, cutoff_quantile),
    ". There are", length(false_sisters),
    "sisterhoods out of", length(unique(sister_d$sister_id)), "will be reomved.")
```

```{r}
table(table(sister_d$sister_id[which(sister_d$true_sisters)]))
table(table(sister_d$sister_id[which(!sister_d$true_sisters)]))
```

### 1. Plot random sisters

Relabel sisters with the info of sample

```{r}
merge <- readRDS(paste0(output_dir, "/all_bt_sample_merge_filtered_gene.RDS"))
merge$sisters_sample <- paste(merge$sample, merge$sisters, sep = "_")
# filter out false sisters
merge$true_sister <- !merge$sisters_sample %in% false_sisters & !is.na(merge$sisters)
merge$sisters_sample[which(merge$sisters_sample %in% false_sisters)] <- paste0(
  merge$sisters_sample[which(merge$sisters_sample %in% false_sisters)],
  "_NA"
)
```

```{r}
all_sister_pairs <- unique(merge$sisters_sample[merge$true_sister])
cat("There are totally", length(all_sister_pairs),
    "unique sisterhoods across all BT samples after filtering.")
```

```{r}
saveRDS(merge, file = paste0(output_dir, "/all_bt_sample_merge_filtered_gene_sisters.RDS"))
```

Randomly picking 20 sister groups to plot

```{r}
set.seed(1)
n <- 10
groups <- sample(all_sister_pairs, n)
highlight_group <- vector("list", length(groups))
names(highlight_group) <- groups
for (i in groups){
  highlight_group[[i]] <- colnames(merge)[which(merge$sisters_sample == i)]
}

# myCol <- ggsci::pal_aaas()(n)
myCol <- c('#88CCEE', '#44AA99', '#117733', '#332288', '#DDCC77',
           '#999933', '#CC6677', '#882255', '#AA4499', "#000000")
names(myCol) <- groups
pca <- PCAPlot(merge, cells.highlight = highlight_group,
               cols.highlight = myCol[groups], sizes.highlight = 4) + NoLegend()


umap <- UMAPPlot(merge, cells.highlight = highlight_group,
               cols.highlight = myCol[groups], sizes.highlight = 4) + NoLegend()

tsne <- TSNEPlot(merge, cells.highlight = highlight_group,
               cols.highlight = myCol[groups], sizes.highlight = 4) + NoLegend()

pca
umap
tsne
```

```{r}
  ggsave(paste0(output_dir, "/highlight_sister_group_",paste0(groups, collapse = "."), "_filtered_sisters_pca.pdf"), plot = pca, device = "pdf", width = 7)
  ggsave(paste0(output_dir, "/highlight_sister_group_",paste0(groups, collapse = "."), "_filtered_sisters_umap.pdf"), plot = umap, device = "pdf", width = 7)
  ggsave(paste0(output_dir, "/highlight_sister_group_",paste0(groups, collapse = "."), "_filtered_sisters_tsne.pdf"), plot = tsne, device = "pdf", width = 7)
```

```{r}
rm(list = ls())
```

# 3. Cell cycle of all BT samples

```{r}
output.dir <- "../data/output/KURAMOCHI_before_treatment/cell_cycle/"

if (!dir.exists(output.dir)) {
  dir.create(output.dir)
}

# Load data

merge <- readRDS("../data/output/KURAMOCHI_before_treatment/all_bt_sample_merge_filtered_gene_sisters.RDS")
merge <- NormalizeData(merge)
merge <- FindVariableFeatures(merge)
merge <- ScaleData(merge, features = rownames(merge))
merge <- RunPCA(merge)

# Cell cycle scoring

# A list of cell cycle markers, from Tirosh et al, 2015, is loaded with Seurat. We can
# segregate this list into markers of G2/M phase and markers of S phase
s.genes <- merge@misc$gene_meta$ensembl[which(merge@misc$gene_meta$symbol %in% cc.genes$s.genes)]
g2m.genes <- merge@misc$gene_meta$ensembl[which(merge@misc$gene_meta$symbol %in% cc.genes$g2m.genes)]
merge <- CellCycleScoring(merge, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)

merge <- RunPCA(merge, features = c(s.genes, g2m.genes))
```
```{r}
DimPlot(merge)
ggsave(paste0(output.dir, "all_bt_cell_cycle.png"))

DimPlot(merge, split.by = "sample")
ggsave(paste0(output.dir, "all_bt_cell_cycle_split_by_sample.png"),
       width = 20)

samples <- c("NK1", "NK2", "Carbo1", "Carbo2", "Olaparib1", "Olaparib2")
plots <- list()
for (s in samples) {
  tmp <- subset(
    merge,
    cells = rownames(merge@meta.data)[merge$sample == s]
  )
  DimPlot(
    tmp,
    split.by = "drugSens",
    group.by = "Phase"
  ) +
    labs(
      title = s
    )
  ggsave(
    paste0(output.dir, s, "_bt_cell_cycle_split_by_drug_sensitivity.png"),
    width = 10, height = 5)
  
  plots[[s]] <- ggplot(
    tmp@meta.data %>% 
      rownames_to_column("id"),
    aes(fill=Phase, x=drugSens)
  ) + 
    geom_bar(position="fill", stat = "count") +
    labs(
      title = s,
      y = "Proportion",
      x = ""
    ) +
    theme_classic() +
    theme(
      plot.title = element_text(hjust = 0.5)
    )
}
p <- ggarrange(plotlist = plots, ncol = 2, nrow = 3)
ggsave(
  paste0(output.dir, "cell_cyle_drug_sensitivity_proportion.png"),
  p, width = 10, height = 12)

# Validated genes

validated_genes <- merge@misc$gene_meta[which(
  merge@misc$gene_meta$symbol %in% c(
    "HMGB1", "PRKDC", "XRCC5", "HNRNPA1", "NPM1", "HNRNPA3", "MT-CO1", "MT-CO3",
  "MT-ND1", "MT-RNR2", "MDK", "FIS1")),]

plots <- list()
for (i in 1:nrow(validated_genes)) {
  plots[[i]] <- RidgePlot(
    merge,
    features = validated_genes$ensembl[i],
  ) +
    ggtitle(validated_genes$symbol[i]) +
    theme(
      plot.title = element_text(hjust = 0.5),
      axis.title.y = element_blank()
    ) +
    NoLegend()
}

p <- ggarrange(plotlist = plots, ncol = 2, nrow = 6)
ggsave(paste0(output.dir, "validated_genes_expression_distribution.png"),
       width = 10, height = 15)
```

# 4. (Batch job) DEG Pre-resistant VS Pre-sensitive

The DEGs are calculated via batch job in 4 different settings:

# 5. (Batch job) Sister conserve genes

This part is calculated with batch job. The codes are in: "./calculate_conserve_gene.R" and "./calculate_conserve_gene.sh"

# 6. scVelo

prepare files required by python code

```{r}
rm(list = ls())
library(Seurat)
options(bitmapType='cairo')
output_dir <- "../data/output/scvelo/"
if (!dir.exists(output_dir)){
  dir.create(output_dir, recursive = TRUE)
}
merge <- readRDS("../data/output/KURAMOCHI_before_treatment/all_bt_sample_merge_filtered_gene_sisters.RDS")
```

```{r}
cells = colnames(merge)
cells <- sub("-1", "", cells)
cells <- sub("BT_", "", cells)
cells <- sub("_", ":", cells)
cells <- sub("_", "-", cells)
write.csv(cells, paste0(output_dir, "cells.csv"), row.names = FALSE)
genes <- rownames(merge)
write.csv(genes, paste0(output_dir, "genes.csv"), row.names = FALSE)

# Sister cell
sisters <- data.frame(sisters = merge$sisters,
                      cell = cells,
                      stringsAsFactors = FALSE)
sisters$sisters[is.na(sisters$sisters)] <- 0
write.csv(sisters, paste0(output_dir, "sisters.csv"), row.names = FALSE)
# Drug sensitive
sen <- data.frame(drugSens = paste(merge$sample, merge$drugSens),
                  cell = cells,
                  stringsAsFactors = FALSE)
sen$drugSens <- gsub("NA", "no label", sen$drugSens)
write.csv(sen, paste0(output_dir, "DrugSens.csv"), row.names = FALSE)

write.csv(merge@misc$gene_meta, paste0(output_dir, "gene_meta.csv"), row.names = FALSE)
```

RNA velocity analysis was run on cluster server. The scripts are "runscvelo.py", "runscvelo.sh"

```{r}
sessionInfo()
```
