---
title: "Drug induced signature scRNA data"
output: html_notebook
---

```{r}
library(tidyverse)
library(Seurat)
library(PharmacoGx)
sessionInfo()
```

```{r}
output_dir <- "../data/output/"
```

# Load data

```{r}
qc_statistic <- NULL
merge <- NULL
treatments <- c(
  "Bortezomib","DMSO", "Ixazomib_citrate", "Pevonedistat_hydrochloride_L",
  "Carfilzomib", "Delanzomib", "Nocodazole", "Pyrrolidinedithiocarbamate_ammonium",
  "Clofarabine", "GW843682X", "Pevonedistat_hydrochloride_H", "Sitagliptin"
)
for (r in 1:2) {
  for (t in treatments) {
    data <- Read10X(
      data.dir = paste0("../data/sample_", r, "/per_sample_outs/", t, "/count/sample_filtered_feature_bc_matrix/"),
      gene.column = 1 # Use Ensembl ID as the row names
    )
    data <- CreateSeuratObject(counts = data$`Gene Expression`, project = paste0("replicate_", r), min.cells = 3)
    data$treatment <- t
    tmp <- data.frame(
      gene = nrow(data),
      cell_before_qc = ncol(data),
      treatment = t,
      replicate = paste0("replicate_", r)
    )

    if (!is.null(qc_statistic)){
      qc_statistic <- rbind.data.frame(qc_statistic, tmp)
    } else {
      qc_statistic <- tmp
    }
    
    if (!is.null(merge)) {
      merge <- merge(merge, data)
    } else {
      merge <- data
    }
  }
}

 # add gene ID symbol exchange table to "misc" slot
genes <- read.table("../data/sample_1/per_sample_outs/Bortezomib/count/sample_filtered_feature_bc_matrix/features.tsv.gz", stringsAsFactors = FALSE) %>% 
  select(1, 2) %>% 
  rename(ensembl = V1, symbol = V2)
genes$ensembl <- sub("_", "-", genes$ensembl)
merge@misc[['gene_meta']] <- genes
```

# Quality control

UMI count

**Please adjust the values of 'low.count' and 'up.count' in the following code chuck.**

```{r}
low.count <- 10000 # Adjust the lower bounder of UMI count here
up.count <- 100000 # Adjust the upper bounder of UMI count here
ggplot(data = merge[["nCount_RNA"]], aes(nCount_RNA)) + 
  geom_histogram(aes(y = ..density..), binwidth = 200) +
  geom_vline(aes(xintercept = up.count), colour="red") +
  geom_vline(aes(xintercept = low.count), colour="red")
```

Feature count

**Please adjust the values of 'low.feature' and 'up.feature' in the following code chuck.**

```{r}
low.feature <- 3000 # Adjust the lower bounder of Feature count here
up.feature <- 12500 # Adjust the upper bounder of Feature count here
ggplot(data = merge[["nFeature_RNA"]], aes(nFeature_RNA)) + 
  geom_histogram(aes(y = ..density..), binwidth = 100) +
  geom_vline(aes(xintercept = up.feature), colour="red") +
  geom_vline(aes(xintercept = low.feature), colour="red")
```

Percentage of mitochondrial RNA

**Please adjust the value of 'up.mt' in the following code chuck.**

```{r}
merge[["percent.mt"]] <- PercentageFeatureSet(merge, features = merge@misc[['gene_meta']]$ensembl[grepl("^MT-", merge@misc[['gene_meta']]$symbol)])
up.mt <- 20 # Adjust the lower bounder of percentage of mitochondrial RNA here
ggplot(data = merge[["percent.mt"]], aes(percent.mt)) + 
  geom_histogram(aes(y = ..density..), binwidth = 1) +
  geom_vline(aes(xintercept = up.mt), colour="red")
```

All the thresholds used here are set in the previous session.

```{r}
merge_f <- subset(merge, subset = nCount_RNA > low.count & nCount_RNA < up.count)
merge_f <- subset(merge_f, subset = nFeature_RNA > low.feature & nFeature_RNA < up.feature)
merge_f <- subset(merge_f, subset = percent.mt < up.mt)
```

```{r}
tmp <- data.frame(table(merge_f$treatment, merge_f$orig.ident))
names(tmp) <- c("treatment", "replicate", "cell_after_qc")
QC_statistic <- left_join(qc_statistic, tmp, by = c("treatment", "replicate")) %>% 
  select(treatment, replicate, everything())

write.csv(QC_statistic, "../data/output/cell_gene_number_before_after_qc.csv", row.names = FALSE)
```

# Dimension reduction

```{r}
merge <- NormalizeData(merge, normalization.method = "LogNormalize", scale.factor = 10000)
merge <- FindVariableFeatures(merge, selection.method = "vst", nfeatures = 2000)
all.genes <- rownames(merge)
merge <- ScaleData(merge, features = all.genes)
merge <- RunPCA(merge, features = VariableFeatures(object = merge))
merge <- JackStraw(merge, num.replicate = 100, dims = 50)
merge <- ScoreJackStraw(merge, dims = 1:50)
ElbowPlot(merge, ndims = 50)
```

```{r}
merge <- RunUMAP(merge, dims = 1:30)
DimPlot(merge, reduction = "umap")
ggsave("../data/output/umap_group_by_replicate.png")
```

```{r}
DimPlot(merge, reduction = "umap", split.by = "treatment")
ggsave("../data/output/umap_group_by_treatment.png", width = 20, height = 4)
```

```{r}
saveRDS(merge, "../data/output/drug_induced_sample_all_merged.RDS")
```

# DEG

```{r}
Idents(merge) <- "treatment"
treatments <- c(
  "DMSO", "Bortezomib","Ixazomib_citrate", "Pevonedistat_hydrochloride_L",
  "Carfilzomib", "Delanzomib", "Nocodazole", "Pyrrolidinedithiocarbamate_ammonium",
  "Clofarabine", "GW843682X", "Pevonedistat_hydrochloride_H", "Sitagliptin"
)
for (t in setdiff(treatments, "DMSO")) {
  deg <- NULL
  for (r in 1:2){
    deg_tmp <- FindMarkers(
      subset(merge, cells = colnames(merge)[merge$orig.ident == paste0("replicate_", r)]),
      ident.1 = t,
      ident.2 = "DMSO",
      test.use = "wilcox",
      logfc.threshold = 0,
      min.pct = 0,
      verbose = TRUE
      )
    deg_tmp <- deg_tmp %>%
      rownames_to_column(var = "ensembl") %>%
      left_join(as.data.frame(merge@misc$gene_meta), by = "ensembl")
    
    write.csv(deg_tmp, paste0("../data/output/deg_", t, "_replicate_", r, "_DMSO_wilcox.csv"), row.names = FALSE)
    if (r == 1 ){
      colnames(deg_tmp)[!colnames(deg_tmp) %in% c("ensembl", "symbol")] <- paste0(t, "_1_", colnames(deg_tmp)[!colnames(deg_tmp) %in% c("ensembl", "symbol")])
      deg <- deg_tmp
    } else {
      colnames(deg_tmp)[!colnames(deg_tmp) %in% c("ensembl", "symbol")] <- paste0(t, "_2_", colnames(deg_tmp)[!colnames(deg_tmp) %in% c("ensembl", "symbol")])
      deg <- deg %>% 
        left_join(deg_tmp, by = c("ensembl", "symbol")) %>% 
        select(ensembl, symbol, everything())
    }
  }
  write.csv(deg, paste0("../data/output/deg_", t, "_merge_DMSO_wilcox.csv"), row.names = FALSE)
}
```

# Connectivity score

## Merge all log2FC

```{r}
preR_preS <- read.csv("../../ReSisTrace_manuscript/data/output/Figure4/Figure4B_all_signatures_logFC_p.csv")
```

```{r}
log2FC_drug_sc <- NULL
for (t in setdiff(treatments, "DMSO")){
  tmp <- read.csv(paste0("../data/output/deg_", t, "_merge_DMSO_wilcox.csv"))
  tmp[[paste0(t, "_DMSO_log2FC")]] <- (tmp[[paste0(t, "_1_avg_log2FC")]] + tmp[[paste0(t, "_2_avg_log2FC")]])/2
  tmp[[paste0(t, "_DMSO_p")]] <- (tmp[[paste0(t, "_1_p_val")]] + tmp[[paste0(t, "_2_p_val")]])/2
  tmp[[paste0(t, "_DMSO_p_adj")]] <- (tmp[[paste0(t, "_1_p_val_adj")]] + tmp[[paste0(t, "_2_p_val_adj")]])/2
  tmp <- dplyr::select(tmp,
    ensembl, symbol, paste0(t, "_DMSO_log2FC"), 
    paste0(t, "_DMSO_p"), paste0(t, "_DMSO_p_adj")
  )
  if (!is.null(log2FC_drug_sc)) {
    log2FC_drug_sc <- full_join(log2FC_drug_sc, tmp, by = c("ensembl", "symbol"))
  } else {
    log2FC_drug_sc <- tmp
  }
}
```

```{r}
data_merge <- preR_preS %>%
  left_join(log2FC_drug_sc, by = c("ensembl", "symbol")) %>% 
  select(-X) %>% 
  filter(conserve_gene == TRUE)

write.csv(data_merge, "../data/output/preR_preS_drug_signature_all.csv", row.names = FALSE)
```

## Calculate connectivity scores

```{r include=FALSE}
drugs <- c(
  "Bortezomib","Ixazomib_citrate", "Pevonedistat_hydrochloride_L",
  "Carfilzomib", "Delanzomib", "Nocodazole", "Pyrrolidinedithiocarbamate_ammonium",
  "Clofarabine", "GW843682X", "Pevonedistat_hydrochloride_H", "Sitagliptin"
)
treatments <- c("Carbo", "NK", "Olaparib")

data_merge <- preR_preS %>%
  left_join(log2FC_drug_sc, by = c("ensembl", "symbol")) %>% 
  select(-X) %>% 
  filter(conserve_gene == TRUE)

for (t in treatments) {
  df <- NULL
  for (i in drugs){
    set.seed(123)
    drug_signature <- data_merge[[paste0(i, "_DMSO_log2FC")]]
    names(drug_signature) <- data_merge$symbol
    drug_signature <- na.omit(drug_signature)
    
    # preR-preS adjusted growth control
    preR_signature <- data_merge[[paste0(t, "_log2FC_preR_preS_adj")]]
    names(preR_signature) <- data_merge$symbol
    preR_signature <- na.omit(preR_signature)
    
    connect <- connectivityScore(
      x = drug_signature,
      y = preR_signature,
      method = "fgsea",
      nperm = 1000
    )
    
    tmp <- data.frame(
      drug = i,
      connect_score_log2FC_preR_preS_adj = connect[1],
      connect_p_log2FC_preR_preS_adj = connect[2],
      pearson_n_log2FC_preR_preS_adj = length(preR_signature)
    )
    
    # preR-preS
    
    preR_signature <- data_merge[[paste0(t, "_log2FC_preR_preS")]]
    names(preR_signature) <- data_merge$symbol
    preR_signature <- na.omit(preR_signature)
    
    connect <- connectivityScore(
      x = drug_signature,
      y = preR_signature,
      method = "fgsea",
      nperm = 1000
    )
   
    tmp$connectivity_score_log2FC_preR_preS = connect[1]
    tmp$connectivity_p_log2FC_preR_preS = connect[2]
    tmp$connectivity_n_log2FC_preR_preS = length(preR_signature)
    
    # filter by bootstrap p < 0.05
    data_tmp <- data_merge[data_merge[[paste0(t, "_boot_p")]] < 0.05, ]
    
    drug_signature <- data_tmp[[paste0(i, "_DMSO_log2FC")]]
    names(drug_signature) <- data_tmp$symbol
    drug_signature <- na.omit(drug_signature)
    
    # preR-preS adjusted growth control
    
    preR_signature <- data_tmp[[paste0(t, "_log2FC_preR_preS_adj")]]
    names(preR_signature) <- data_tmp$symbol
    preR_signature <- na.omit(preR_signature)
    
    connect <- connectivityScore(
      x = drug_signature,
      y = preR_signature,
      method = "fgsea",
      nperm = 1000
    )
    
    tmp$connectivity_score_log2FC_preR_preS_adj_boot_p = connect[1]
    tmp$connectivity_p_log2FC_preR_preS_adj_boot_p = connect[2]
    tmp$connectivity_n_log2FC_preR_preS_adj_boot_p = length(preR_signature)
    
    # filter by permutation p < 0.05
    data_tmp <- data_merge[data_merge[[paste0(t, "_perm_p")]] < 0.05, ]
    
    drug_signature <- data_tmp[[paste0(i, "_DMSO_log2FC")]]
    names(drug_signature) <- data_tmp$symbol
    drug_signature <- na.omit(drug_signature)
    
    # preR-preS adjusted growth control
    
    preR_signature <- data_tmp[[paste0(t, "_log2FC_preR_preS_adj")]]
    names(preR_signature) <- data_tmp$symbol
    preR_signature <- na.omit(preR_signature)
    
    connect <- connectivityScore(
      x = drug_signature,
      y = preR_signature,
      method = "fgsea",
      nperm = 1000
    )
    
    tmp$connectivity_score_log2FC_preR_preS_adj_perm_p = connect[1]
    tmp$connectivity_p_log2FC_preR_preS_adj_perm_p = connect[2]
    tmp$connectivity_n_log2FC_preR_preS_adj_perm_p = length(preR_signature)

    # filter by permutation p < 0.05 and bootstrap p < 0.05
    data_tmp <- data_merge[
      data_merge[[paste0(t, "_perm_p")]] < 0.05 &
        data_merge[[paste0(t, "_boot_p")]] < 0.05,
      ]
    
    drug_signature <- data_tmp[[paste0(i, "_DMSO_log2FC")]]
    names(drug_signature) <- data_tmp$symbol
    drug_signature <- na.omit(drug_signature)
    
    # preR-preS adjusted growth control
    
    preR_signature <- data_tmp[[paste0(t, "_log2FC_preR_preS_adj")]]
    names(preR_signature) <- data_tmp$symbol
    preR_signature <- na.omit(preR_signature)
    
    connect <- connectivityScore(
      x = drug_signature,
      y = preR_signature,
      method = "fgsea",
      nperm = 1000
    )
    
    tmp$connectivity_score_log2FC_preR_preS_adj_perm_boot_p = connect[1]
    tmp$connectivity_p_log2FC_preR_preS_adj_perm_boot_p = connect[2]
    tmp$connectivity_n_log2FC_preR_preS_adj_perm_boot_p = length(preR_signature)
    
    # filter by preR-preS adjusted p < 0.05
    data_tmp <- data_merge[data_merge[[paste0(t, "_preR_preS_p_adj")]] < 0.05, ]
    
    drug_signature <- data_tmp[[paste0(i, "_DMSO_log2FC")]]
    names(drug_signature) <- data_tmp$symbol
    drug_signature <- na.omit(drug_signature)
    
    # preR-preS adjusted growth control
    
    preR_signature <- data_tmp[[paste0(t, "_log2FC_preR_preS")]]
    names(preR_signature) <- data_tmp$symbol
    preR_signature <- na.omit(preR_signature)
    
    connect <- connectivityScore(
      x = drug_signature,
      y = preR_signature,
      method = "fgsea",
      nperm = 1000
    )
    
    tmp$connectivity_score_log2FC_preR_preS_p_adj = connect[1]
    tmp$connectivity_p_log2FC_preR_preS_p_adj = connect[2]
    tmp$connectivity_n_log2FC_preR_preS_p_adj = length(preR_signature)
    
    # filter by preR-preS p < 0.05
    data_tmp <- data_merge[data_merge[[paste0(t, "_preR_preS_p")]] < 0.05, ]
    drug_signature <- data_tmp[[paste0(i, "_DMSO_log2FC")]]
    names(drug_signature) <- data_tmp$symbol
    drug_signature <- na.omit(drug_signature)
    
    # preR-preS adjusted growth control
    
    preR_signature <- data_tmp[[paste0(t, "_log2FC_preR_preS")]]
    names(preR_signature) <- data_tmp$symbol
    preR_signature <- na.omit(preR_signature)
    
    connect <- connectivityScore(
      x = drug_signature,
      y = preR_signature,
      method = "fgsea",
      nperm = 1000
    )
    
    tmp$connectivity_score_log2FC_preR_preS_p = connect[1]
    tmp$connectivity_p_log2FC_preR_preS_p = connect[2]
    tmp$connectivity_n_log2FC_preR_preS_p = length(preR_signature)
    
    # posT-preT
    drug_signature <- data_merge[[paste0(i, "_DMSO_log2FC")]]
    names(drug_signature) <- data_merge$symbol
    drug_signature <- na.omit(drug_signature)
    
    preR_signature <- data_merge[[paste0(t, "_log2FC_posT_preT")]]
    names(preR_signature) <- data_merge$symbol
    preR_signature <- na.omit(preR_signature)
    
    connect <- connectivityScore(
      x = drug_signature,
      y = preR_signature,
      method = "fgsea",
      nperm = 1000
    )

    tmp$connectivity_score_log2FC_posT_preT = connect[1]
    tmp$connectivity_p_log2FC_posT_preT = connect[2]
    tmp$connectivity_n_log2FC_posT_preT = length(preR_signature)
    
    # posT-preT adjusted growth control
    drug_signature <- data_merge[[paste0(i, "_DMSO_log2FC")]]
    names(drug_signature) <- data_merge$symbol
    drug_signature <- na.omit(drug_signature)
    
    preR_signature <- data_merge[[paste0(t, "_log2FC_posT_preT")]]
    names(preR_signature) <- data_merge$symbol
    preR_signature <- na.omit(preR_signature)
    
    connect <- connectivityScore(
      x = drug_signature,
      y = preR_signature,
      method = "fgsea",
      nperm = 1000
    )

    tmp$connectivity_score_log2FC_posT_preT_adj = connect[1]
    tmp$connectivity_p_log2FC_posT_preT_adj = connect[2]
    tmp$connectivity_n_log2FC_posT_preT_adj = length(preR_signature)
    
    # filter by posT-preT adjusted p < 0.05
    data_tmp <- data_merge[data_merge[[paste0(t, "_posT_preT_p_adj")]] < 0.05, ]
    drug_signature <- data_tmp[[paste0(i, "_DMSO_log2FC")]]
    names(drug_signature) <- data_tmp$symbol
    drug_signature <- na.omit(drug_signature)
    
    # preR-preS adjusted growth control
    
    preR_signature <- data_tmp[[paste0(t, "_log2FC_posT_preT")]]
    names(preR_signature) <- data_tmp$symbol
    preR_signature <- na.omit(preR_signature)
    
    connect <- connectivityScore(
      x = drug_signature,
      y = preR_signature,
      method = "fgsea",
      nperm = 1000
    )
    
    tmp$connectivity_score_log2FC_posT_preT_p_adj = connect[1]
    tmp$connectivity_p_log2FC_posT_preT_p_adj = connect[2]
    tmp$connectivity_n_log2FC_posT_preT_p_adj = length(preR_signature)
    
    # filter by posT-preT p < 0.05
    data_tmp <- data_merge[data_merge[[paste0(t, "_posT_preT_p")]] < 0.05, ]
    drug_signature <- data_tmp[[paste0(i, "_DMSO_log2FC")]]
    names(drug_signature) <- data_tmp$symbol
    drug_signature <- na.omit(drug_signature)
    
    # preR-preS adjusted growth control
    
    preR_signature <- data_tmp[[paste0(t, "_log2FC_posT_preT")]]
    names(preR_signature) <- data_tmp$symbol
    preR_signature <- na.omit(preR_signature)
    
    connect <- connectivityScore(
      x = drug_signature,
      y = preR_signature,
      method = "fgsea",
      nperm = 1000
    )
    tmp$connectivity_score_log2FC_posT_preT_p = connect[1]
    tmp$connectivity_p_log2FC_posT_preT_p = connect[2]
    tmp$connectivity_n_log2FC_posT_preT_p = length(preR_signature)
    
    df <- rbind.data.frame(df, tmp)
  }
  write.csv(
    df,
    paste0(
      output_dir,
      t,
      "_connectivity_score_preR_vs_drug_signature_sister_conserved.csv"
      ),
    row.names = FALSE)
}
```

